---
title: "cost_rp"
author: "Ray Hunter"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=FALSE, message = FALSE, error = FALSE)


#clearing environment
rm(list = ls())

#sourcing data
source("../common.R")
```



### 1) IMPACT OF SITE ACCESSIBILITY AND MATERIALS ON COSTS

## 1.1) SITE ACCESSABILITY: FIND THE NEAREST ROAD TO EACH STREAM REACH and ASSIGN LMH (STOLEN FROM ELJ)
* in common R, the transportation (site accessability) was added to the `flow_stl` data frame


## 1.2) MATERIALS: STREAM ENERGY USED TO CALCULATE MATERIALS COSTS (STOLEN FROM ELJ)
```{r}

flow_stl$slope_perc<-flow_stl$slope*100
flow_stl$slope_degrees <- atan(flow_stl$slope_perc / 100) * (180 / pi)


# Create a function to assign energy values based on slope and wint_width ranges
assign_energy_width <- function(input_df, low_slope_range, medium_slope_range, high_slope_range, low_wint_width_range, medium_wint_width_range, high_wint_width_range, slope) {
  # # Filter for hab_unit == Marsh
  # filtered_df <- input_df %>% filter(hab_unit == "Marsh")
  
  # Create a new column "energy" based on conditions
  filtered_df <- input_df %>%
    mutate(energy = case_when(
      # low slope, low range
      (slope >= low_slope_range[1] & slope <= low_slope_range[2] & 
       wint_width >= low_wint_width_range[1] & wint_width <= low_wint_width_range[2]) |
      # medium slope, low width
      (slope > low_slope_range[2] & slope <= medium_slope_range[2] & 
       wint_width >= low_wint_width_range[1] & wint_width <= low_wint_width_range[2]) |
      # low slope, medium width
      (slope >= low_slope_range[1] & slope <= low_slope_range[2] & 
       wint_width > low_wint_width_range[2] & wint_width <= medium_wint_width_range[2]) 
      ~ "low",
      # low slope, high width
      (slope >= low_slope_range[1] & slope <= low_slope_range[2] & 
       wint_width > medium_wint_width_range[2] & wint_width <= high_wint_width_range[2]) |
      # medium slope, medium width
      (slope > low_slope_range[2] & slope <= medium_slope_range[2] & 
       wint_width > low_wint_width_range[2] & wint_width <= medium_wint_width_range[2]) |
      # high slope, low width
      (slope > medium_slope_range[2] & slope <= high_slope_range[2] & 
       wint_width >= low_wint_width_range[1] & wint_width <= low_wint_width_range[2])
       ~ "medium",
      # high slope, high width
      (slope > medium_slope_range[2] & slope <= high_slope_range[2] & 
       wint_width > medium_wint_width_range[2] & wint_width <= high_wint_width_range[2]) |
      # medium slope, high width
      (slope > low_slope_range[2] & slope <= medium_slope_range[2] & 
       wint_width > medium_wint_width_range[2] & wint_width <= high_wint_width_range[2]) |
      # high slope, medium width
      (slope > medium_slope_range[2] & slope <= high_slope_range[2] & 
       wint_width > low_wint_width_range[2] & wint_width <= medium_wint_width_range[2])
       ~ "high",
      TRUE ~ "Unknown"  # Handle any other cases
    ))
  
  return(filtered_df)
}

# Call the function with the provided dataframe and numeric range bounds
low_slope_range <- c(0, 10)
medium_slope_range <- c(10.0000000001, 20)
high_slope_range <- c(20.0000000001, 1000) #arbitrary large number
low_width_range <- c(0, 20)
medium_width_range <- c(20.0000000001, 40)
high_width_range <- c(40.0000000001, 300) #arbitrary large number

flow_stl <- assign_energy_width(flow_stl, low_slope_range, medium_slope_range, high_slope_range, low_width_range, medium_width_range, high_width_range, slope_degrees)

unique(flow_stl$energy)
```



## 1.3) CALCULATE RELATIVE COST BASED ON MATERIALS (ENERGY) AND SITE ACCESSABILITY (TRANSPORTATION) (STOLEN FROM ELJ)
```{r}
#energy column is called energy (low, medium and high) and transportation is distance_to_nearest_road_mi (values are low, medium and high)

calculate_intermediate_cost <- function(input_df) {
  input_df <- input_df %>%
    mutate(
      intermediate_cost = case_when(
        transportation == "low" & energy == "low" |
        transportation == "low" & energy == "medium" ~ "low",
        transportation == "low" & energy == "high" |
        transportation == "medium" & energy == "low" |
        transportation == "medium" & energy == "medium" ~ "medium",
        transportation == "medium" & energy == "high" |
        transportation == "high" & energy == "low" |
        transportation == "high" & energy == "medium" |
        transportation == "high" & energy == "high" ~ "high",
        # Add more conditions for other combinations here
        TRUE ~ "Unknown"  # Default value if no conditions match
      )
    )
  
  return(input_df)
}
  flow_stl <- calculate_intermediate_cost(flow_stl)

```


#####################################################################################################

### 2) COST OF RIPARIAN PLANTING PROJECTS ($/acre)

```{r}
## matrix with levels of site preparation ( average slope ) and output from materials/accessibility 
# need to measure the average slope around a stream within 50m 
```

#### 2.1) LEVEL OF SITE PREPARATION 

```{r}
                     
# importing elevation raster data
# Data: https://gis.ess.washington.edu/data/raster/tenmeter/byquad/info.html


# listing the dem elevation files 
elev_list <- list.files(path = "/Users/rayhunter/Desktop/GP/gp_anadromites/data/elevation", "dem$", 
                   full.names = TRUE)

#  rasterizing elevation dem files and storing them in a collection 
elev_list_rast <- sprc(lapply(elev_list, rast))

# combining rasters into 1 mosaic
elev_mos <- mosaic(elev_list_rast)

#calculating slope from elevation data
full_slope <- terrain(x = elev_mos, v = "slope", unit = "degrees")
# plot(full_slope)

# projecting NAD83 (which is more accurate and used by the NOAA files) onto raster data (originally using NAD27)  
full_slope <- project(full_slope, "EPSG:26910")

# writeLines(crs(flow_buff))
# writeLines(crs(full_slope))

# creating a 50m buffer around stream edges
flow_buff <- st_buffer(flow_stl, dist = 50, endCapStyle = "FLAT")

# cropping and masking slope raster data to stream buffer
crop_rast <- crop(full_slope, flow_buff, mask = TRUE)
# plot(crop_rast)

#extract mean 
extract <- extract(crop_rast, flow_buff, fun = mean)

#joining to original flow_stl 
flow_stl <- flow_stl %>% 
  mutate(avg_bank_slope = extract$slope)



######## CLASSIFY SLOPE LOW, MED, HIGH ########

low_slope <- c(0, 10)
med_slope <- c(10.001, 20)
high_slope <- c(20.001, Inf)


bank_slope_cost_fun <- function(slope) {
  bank_slope_class <- ifelse(slope > 0 & slope <= 10, "low",
                ifelse(slope > 10 & slope <= 20, "medium", "high"))
  return(bank_slope_class)
}

flow_stl <- flow_stl %>%
  mutate(bank_slope_class = bank_slope_cost_fun(avg_bank_slope),
        #creating acres of riparian vegetaion. may need to be changed
         #2x the area of the stream reach converted to acres
         rip_acres = 2*length*wint_width*0.000247105
         )


```



### 3) COST OF RIPARIAN PLANTING PROJECTS 
### MATERIALS/SITE ACCESSIBILITY X SITE PREPARATION

```{r}
rp_cost <- function(input_df) {
  input_df <- input_df %>%
    mutate(
      #creating a column for low end of cost spectrum 
      lower_cost = rip_acres*case_when(
        intermediate_cost == "low" & bank_slope_class == "low" ~ 5000,
        intermediate_cost == "low" & bank_slope_class == "medium" ~ 20000,
        intermediate_cost == "low" & bank_slope_class == "high" ~ 60000,
        intermediate_cost == "medium" & bank_slope_class == "low" ~ 10000,
        intermediate_cost == "medium" & bank_slope_class == "medium" ~ 45000,
        intermediate_cost == "medium" & bank_slope_class == "high" ~ 70000,
        intermediate_cost == "high" & bank_slope_class == "low" ~ 30000,
        intermediate_cost == "high" & bank_slope_class == "medium" ~ 55000,
        intermediate_cost == "high" & bank_slope_class == "high" ~ 100000,
        # Add more conditions for other combinations here
        TRUE ~ 0  # Default value if no conditions match
      ),
      #creating a column for high end of cost spectrum 
      upper_cost = rip_acres*case_when(
       intermediate_cost == "low" & bank_slope_class == "low" ~ 25000,
        intermediate_cost == "low" & bank_slope_class == "medium" ~ 50000,
        intermediate_cost == "low" & bank_slope_class == "high" ~ 100000,
        intermediate_cost == "medium" & bank_slope_class == "low" ~ 35000,
        intermediate_cost == "medium" & bank_slope_class == "medium" ~ 65000,
        intermediate_cost == "medium" & bank_slope_class == "high" ~ 120000,
        intermediate_cost == "high" & bank_slope_class == "low" ~ 50000,
        intermediate_cost == "high" & bank_slope_class == "medium" ~ 80000,
        intermediate_cost == "high" & bank_slope_class == "high" ~ 135000,
        # Add more conditions for other combinations here
        TRUE ~ 0  # Default value if no conditions match
      )
    )
  
  return(input_df)
}

rp_cost_final <- rp_cost(flow_stl)
```


##SUBBASIN SUMMARY 
```{r}
subba_rp_cost <- rp_cost_final %>% 
  group_by(noaa_subba) %>% 
  summarize(lower_subba_cost = sum(lower_cost),
            upper_subba_cost = sum(upper_cost))
```




#### NEXT STEPS: 
* Problem: all intermediate costs are low b/c transportation is all "low" and energy is all "low" or "medium"
* need to figure out how many acres along each stream need to be restored
  - cost estimates are given in $/acre. Literature says at least 30m inland
  - right now, I multiplied the area of the stream (length x winter width) by 2 and converted to acres. no idea if this is right
  
ASSUMPTIONS
* this assumes full restoration of all banks of every stream. what if we only want to restore partial riparian vegetation?
* assuming riparian vegetation is 2x the area of the stream

















