```{r}
#sourcing data
source("../common.R")

library(sf)
library(rmapshaper)
library(ggplot2)
library(mapview)
```


######################################## Land Use #####################################
#Run this only once
```{r}
######################## making shape files to streamline process ##################### 
# 
# # creating a shape file of the unionized stillguamish basin so we don't have to redo this operation every time
# 
# subs_stl_union <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
#   clean_names() %>%
# st_union() %>%
#   st_as_sf() %>%
#   st_write(here("data", "HARP", "Subbasins_STL", "unionized_basin.shp"))
# 
# 
# 
# #creating a shape file of stilly land use so we dont have to read in all the stillaguamish and skagit data and then transform and filter it each time -->
# 
# land_use_stilly <- land_use %>%
# st_intersection(subs_union) %>%
# st_write(here("data", "land_use", "land_use_stilly.shp"))
```

#READING IN THE ABOVE DATA
```{r}
#reading in unionized stilly subasin shape file -->
subs_union <- st_read(here("data", "HARP", "Subbasins_STL","unionized_basin.shp")) %>% st_transform("EPSG:26910")

#reading in individual stilly subbasins shape file -->
subs_stl <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
  clean_names() %>% 
  select(noaa_subba, shape_area, geometry) %>% 
  st_transform("EPSG:26910") 

#land use parcel data for snohomish and skagit county -->
land_use_stilly <- st_read(here("data", "land_use", "land_use_stilly.shp")) %>% 
  clean_names() %>% select(master_cat, sub_cat, lu_code_id, shape_are, shape_len)

#also read in the target subbasins (benefits to chinook)
subbasins_key <- read_csv(here("data", "land_use", "chinook_ben_subbasins.csv")) %>% select(noaa_subba)

```

# NEW PARCELS LAYER BY SUBBASIN 
```{r}
#creating a new df with parcels by subbasin 

#splitting parcels if they intersect w/ multiple subs 
land_use_subs <- st_intersection(land_use_stilly, subs_stl) %>% 
  #joining attributes of noaa subbasin to parcel data 
  st_join(subs_stl, largest = TRUE) %>% 
  select(master_cat, sub_cat, noaa_subba = noaa_subba.x, subba_area = shape_area.x) %>% 
  mutate(parcel_area_ha = units::drop_units(st_area(.)/10000))



##summary of percent land use in each subbasin (with all values)
summary_land_use <- land_use_subs %>%
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area_ha=sum(parcel_area_ha)) %>% 
  st_drop_geometry() %>% 
  select(-sub_cat) %>% 
  unique() %>%
  # Ungroup the data
  ungroup() %>%
  # Group by 'noaa_subba' again
  group_by(noaa_subba) %>%
  # Calculate the relative percentage for each 'master_cat' within each 'noaa_subba'
  mutate(perc_land_use = round(parcel_area_ha / sum(parcel_area_ha) * 100,2)) %>%
  mutate(perc_land_use = as.numeric(perc_land_use)) %>% 
  mutate(parcel_area_ha = as.numeric(parcel_area_ha)) %>% 
  right_join(subbasins_key)

#summary of percent land use (with values under 1 percent aggregated to Other)
summary_land_use <- land_use_subs %>%
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area_ha = sum(parcel_area_ha)) %>% 
  st_drop_geometry() %>% 
  select(-sub_cat) %>% 
  unique() %>%
  ungroup() %>%
  group_by(noaa_subba) %>%
  mutate(parcel_area_ha = as.numeric(parcel_area_ha)) %>% 
  mutate(perc_land_use = parcel_area_ha / sum(parcel_area_ha) * 100,  # Calculate perc_land_use
         master_cat = case_when(
           perc_land_use < 1 ~ "Other",
           TRUE ~ as.character(master_cat)  # Keep the original master_cat otherwise
         )) %>%
  mutate(perc_land_use = round(perc_land_use, 2)) %>% 
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area_ha = sum(parcel_area_ha)) %>% 
  mutate(perc_land_use = round(sum(perc_land_use), 2)) %>% unique()%>% 
  right_join(subbasins_key)
```

##Land Use Figure
```{r}
#preparing the data for the figure
  #arrange in descending
  #make factors
summary_land_use <- summary_land_use %>% 
  mutate(noaa_subba = fct_reorder(noaa_subba, parcel_area_ha, .fun = sum),
         master_cat = factor(master_cat,
                                   levels = c("Forest Lands", 
                                              "Rural Character Residential",
                                              "Agricultural Area",
                                              "Mineral Resource Area",
                                              "Water",
                                              "ROW",
                                              "Public",
                                              "Intensive Urban",
                                              "Other"))) %>% 
  #cleaning names
  mutate(subbasin_name = str_replace_all(noaa_subba, "_", " "),
         subbasin_name = str_replace_all(subbasin_name, "Stillaguamish", ""),
         subbasin_name = str_replace_all(subbasin_name, "mainstem", "Mainstem")
  ) %>% filter(noaa_subba!="mainstem_South_Fork_Stillaguamish_02") #remove this subbasin from the figure
  

#define custom colors
# color themes 
custom_colors <- c("Agricultural Area" = "#066b6b",
                   "Forest Lands" = "#03045E", 
                   "Mineral Resource Area" = "#b5dee8",
                   "Other" = "#71a1bd",
                   "Public" = "#599e94",
                   "ROW" = "#144040",
                   "Rural Character Residential" = "#28AFB0",
                   "Intensive Urban" = "#a3c2c2",
                   "Water" = "#1d538a" 
                   )

#figure itself
land_use_barchart<- summary_land_use %>% 
  ggplot(aes(
    # x = noaa_subba, 
             x = reorder(subbasin_name, subba_area),
             y = parcel_area_ha/1000, fill = master_cat)) +
  theme_minimal() +
  geom_col() +
  labs(fill = "Land Use Category", 
       x = "", 
       y = "Area (Thousand Hectares)",
       title = "Land Use Category",
       subtitle = "by subbasin",
       caption = "Data Source: 
United States. (2024) U.S. Census of Agriculture.") +  ##CHANGE THIS
  coord_flip() + 
  scale_fill_manual(values = custom_colors) +
  # pushing y axis labels to edhe of data 
  scale_y_continuous(expand = c(0,0)) +
  theme(
    #axes
    axis.text = element_text(size = 10),
    axis.title.x = element_text(margin = margin(t = 12, r = 0, b = 30, l = 20), size = 12),
    #legend
    legend.position = c(0.5, -0.175),
    legend.direction = "horizontal",
    legend.title = element_blank(),
    # legend.box.background = element_rect(color = "grey"))
    # titles
    plot.title = element_text(hjust = 0, vjust = 1.8, size = 16),
    plot.subtitle = element_text(hjust = 0, vjust = 2.2),
    plot.caption = element_text(face = "italic", size = 6),
    # gridlines
    panel.grid.major.y  = element_blank(),
    panel.grid.minor.x  = element_blank()
  ) 

land_use_barchart

ggsave("../final_prods/figures/landuse_barchart.jpeg", plot = land_use_barchart, device = "jpeg", width = 10, height = 7, units = "in")

```


## Mapping land use
```{r}
ggplot() + 
  geom_sf(data = land_use_subs, aes(fill = master_cat))
```

#################################### Floodplain ###################################
## Creating shapefiles to correspond to the newly created floodplain habitat adjacent to the river

```{r}
#want to create a buffer around flow_stl
bankfull_stl <- st_buffer(flow_stl, dist = flow_stl$bf_width)

#want to create a buffer around the buffer 1/3 * bf_width
floodplain_hab_stl_river<-st_buffer(bankfull_stl, dist = flow_stl$bf_width*(1/3))

st_write(floodplain_hab_stl_river, here("data", "fp_shapefiles", "sidechannel_shapefiles.shp"), append = FALSE)
```

#combine new floodplain shapefile with old floodplain shapefile
```{r}
#changing crs to match floodplain habitat crs
subs_mod <- subs_stl %>% 
  select(noaa_subba) %>% 
  st_transform(26910) 

#filter floodplain to only include historic
hist_flood_stl <- st_join(subs_mod, filter(flood_stl, period %in% "Hist"), join=st_intersects)

#also filter by floodplain habitat with benefit to chinook
flood_ben<-stilly_bens %>% filter(scenario=="Floodplain" &pop == "fall_chinook" & perc_change > 0)  

#filter flood_ben to hist_flood_stl to only get fp with benefits to chinook
hist_flood_stl<-hist_flood_stl %>% filter(noaa_subba %in% unique(flood_ben$subbasin))

# hist_flood_stl<-left_join(flood_ben, hist_flood_stl, by=c("subbasin"="noaa_subba"))
hist_flood_stl<-sf_object <- st_as_sf(hist_flood_stl)

```

#overlap land use and floodplain shapefile
CHECK THIS ON A MAP
```{r}
# #overlap the land use with the floodplain_hab_stl
# farm_stl<-land_use_subs %>% filter(master_cat=="Agricultural Area")
# floodplain_land_use <- st_join(farm_stl, floodplain_hab_stl, join = st_intersects)
# # floodplain_land_use <- st_intersection(farm_stl, floodplain_hab_stl)
```


### Getting a cost per subbasin value
```{r}
# historical floodplain (ponds and marshes)
flood_hist <- flood_stl %>% 
  filter(period == "Hist") %>% 
  select(hab_unit, geometry)

# side channel habitat
flood_sc <- floodplain_hab_stl_river %>% 
  # select(hab_unit, geometry) %>% 
  st_union() %>% 
  st_sf() %>% 
  # naming it side channel
   mutate(hab_unit = "side_channel") 

# combine the two habitats into 1 df
comb_flood <- rbind(flood_hist, flood_sc) 

#finding just the ag land from the land class
ag_land <- land_use_subs %>% 
  filter(master_cat == "Agricultural Area") %>% 
  # unionize parcels
  st_union()

# finding floodplain area on ag land
flood_ag <-  st_intersection(comb_flood, ag_land) 

# finding floodplain ag land within each subbasin
flood_ag_subba <- st_intersection(flood_ag, subs_stl) %>%
  select(-shape_area) %>%
  # finding area in m2, HA, and acres
  mutate(area_m2 = units::drop_units(st_area(.)),
         area_ha = area_m2*0.0001,
         area_acre = area_m2*0.000247105, 
         cost = area_acre*22374) %>%
  inner_join(subbasins_key) #only keeping subbasins that have benefits to chinook

# viewing on map 
mapview(ag_land) +
mapview(flood_ag_subba, zcol = "hab_unit")

#making a summarized table with the values
flood_ag_subba_sum<-flood_ag_subba %>% group_by(noaa_subba, hab_unit) %>% mutate(total_subba_cost=sum(cost), total_subba_area_ha=sum(area_ha)) %>% st_drop_geometry() %>% 
  select(noaa_subba, total_subba_area_ha, total_subba_cost, cost, hab_unit) %>% unique()

# #by subbasin alone
# flood_ag_subba_sum<-flood_ag_subba %>% group_by(noaa_subba) %>% mutate(total_subba_cost=sum(cost), total_subba_area_acres=sum(area_acre)) %>% st_drop_geometry() %>% 
#   select(noaa_subba, total_subba_area_acres, total_subba_cost) %>% unique()

```


##making a figure showing the costs per subbasin


```{r}
#preparing the data for the figure
  #arrange in descending
  #make factors
flood_ag_subba_sum <- flood_ag_subba_sum %>% 
  mutate(noaa_subba = fct_reorder(noaa_subba, total_subba_cost, .fun = sum),
         master_cat = factor(hab_unit,
                                   levels = c("side_channel","Marsh","Pond"))) %>% 
  #cleaning names
  mutate(subbasin_name = str_replace_all(noaa_subba, "_", " "),
         subbasin_name = str_replace_all(subbasin_name, "Stillaguamish", ""),
         subbasin_name = str_replace_all(subbasin_name, "mainstem", "Mainstem"),
         hab_unit = str_replace_all(hab_unit, "side_channel", "Side Channel")
  ) 

#define custom colors
# color themes 
custom_colors <- c("Pond" = "#19647E",
                   # "#1F271B",
                   "Marsh" = "#03045E", 
                   "Side Channel" = "#28AFB0"
                   )



#figure itself
ag_land_price_barchart<- flood_ag_subba_sum %>% 
  ggplot(aes(
    # x = noaa_subba, 
             x = reorder(subbasin_name, total_subba_cost),
             y = total_subba_cost/1000000, fill = hab_unit)) +
  theme_minimal() +
  geom_col() +
  labs(fill = "Habitat Type", 
       x = "", 
       y = "Cost (Million $)",
       title = "Price of Agricultural Floodplain Habitat",
       subtitle = "by subbasin",
       caption = "Data Source: 
United States. (2024) U.S. Census of Agriculture.") +  ##CHANGE THIS
  coord_flip() + 
  scale_fill_manual(values = custom_colors) +
  # pushing y axis labels to edhe of data 
  scale_y_continuous(expand = c(0,0)) +
  theme(
    #axes
    axis.text = element_text(size = 10),
    axis.title.x = element_text(margin = margin(t = 12, r = 0, b = 30, l = 20), size = 12),
    #legend
    legend.position = c(0.5, -0.175),
    legend.direction = "horizontal",
    legend.title = element_blank(),
    # legend.box.background = element_rect(color = "grey"))
    # titles
    plot.title = element_text(hjust = 0, vjust = 1.8, size = 16),
    plot.subtitle = element_text(hjust = 0, vjust = 2.2),
    plot.caption = element_text(face = "italic", size = 6),
    # gridlines
    panel.grid.major.y  = element_blank(),
    panel.grid.minor.x  = element_blank()
  ) 

ag_land_price_barchart

ggsave("../final_prods/figures/ag_land_price_barchart.jpeg", plot = ag_land_price_barchart, device = "jpeg", width = 10, height = 7, units = "in")
```


##overall numbers of land use (entire Stillaguamish level) and overall numbers of land type in the subbasin overall
```{r}
#all acreage (not just overlapping with floodplain)
stilly_land_use_all<-summary_land_use %>% group_by(master_cat) %>% 
  mutate(land_use_area_stilly_ha=sum(parcel_area_ha)) %>% select(master_cat, land_use_area_stilly_ha) %>% unique()


##just the ag land that overlaps with FP
stilly_land_use_fp_overlap<-flood_ag_subba_sum %>% select(noaa_subba, total_subba_area_ha, total_subba_cost) %>% unique()
sum(stilly_land_use_fp_overlap$total_subba_area_ha)
sum(stilly_land_use_fp_overlap$total_subba_cost)
```








