```{r}
#sourcing data
source("../common.R")

library(sf)
library(rmapshaper)
library(ggplot2)
```


######################################## Land Use #####################################
#Run this only once
```{r}
######################## making shape files to streamline process ##################### 
# 
# # creating a shape file of the unionized stillguamish basin so we don't have to redo this operation every time
# 
# subs_stl_union <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
#   clean_names() %>%
# st_union() %>%
#   st_as_sf() %>%
#   st_write(here("data", "HARP", "Subbasins_STL", "unionized_basin.shp"))
# 
# 
# 
# #creating a shape file of stilly land use so we dont have to read in all the stillaguamish and skagit data and then transform and filter it each time -->
# 
# land_use_stilly <- land_use %>%
# st_intersection(subs_union) %>%
# st_write(here("data", "land_use", "land_use_stilly.shp"))
```

#READING IN THE ABOVE DATA
```{r}
#reading in unionized stilly subasin shape file -->
subs_union <- st_read(here("data", "HARP", "Subbasins_STL", "unionized_basin.shp")) %>% 
  st_transform("EPSG:26910")

#reading in individual stilly subbasins shape file -->
subs_stl <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
  clean_names() %>% 
  select(noaa_subba, shape_area, geometry) %>% 
  st_transform("EPSG:26910") 

#land use parcel data for snohomish and skagit county -->
land_use_stilly <- st_read(here("data", "land_use", "land_use_stilly.shp")) %>% 
  clean_names() %>% select(master_cat, sub_cat, lu_code_id, shape_are, shape_len)

#also read in the target subbasins (benefits to chinook)
subbasins_key<-read_csv(here("data", "land_use", "chinook_ben_subbasins.csv")) %>% select(noaa_subba)
```

#NEW PARCELS LAYER BY SUBBASIN 
```{r}
#creating a new df with parcels by subbasin 

#splitting parcels if they intersect w/ multiple subs 
land_use_subs <- st_intersection(land_use_stilly, subs_stl) %>% 
  #joining attributes of noaa subbasin to parcel data 
  st_join(subs_stl, largest = TRUE) %>% 
  select(master_cat, sub_cat, noaa_subba = noaa_subba.x, subba_area = shape_area.x) %>% 
  mutate(parcel_area = st_area(.))

##summary of percent land use in each subbasin (with all values)
summary_land_use <- land_use_subs %>%
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area=sum(parcel_area)) %>% 
  st_drop_geometry() %>% 
  select(-sub_cat) %>% 
  unique() %>%
  # Ungroup the data
  ungroup() %>%
  # Group by 'noaa_subba' again
  group_by(noaa_subba) %>%
  # Calculate the relative percentage for each 'master_cat' within each 'noaa_subba'
  mutate(perc_land_use = round(parcel_area / sum(parcel_area) * 100,2)) %>%
  mutate(perc_land_use = as.numeric(perc_land_use)) %>% 
  mutate(parcel_area = as.numeric(parcel_area)) %>% 
  right_join(subbasins_key)

#summary of percent land use (with values under 1 percent aggregated to Other)
summary_land_use <- land_use_subs %>%
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area = sum(parcel_area)) %>% 
  st_drop_geometry() %>% 
  select(-sub_cat) %>% 
  unique() %>%
  ungroup() %>%
  group_by(noaa_subba) %>%
  mutate(parcel_area = as.numeric(parcel_area)) %>% 
  mutate(perc_land_use = parcel_area / sum(parcel_area) * 100,  # Calculate perc_land_use
         master_cat = case_when(
           perc_land_use < 1 ~ "Other",
           TRUE ~ as.character(master_cat)  # Keep the original master_cat otherwise
         )) %>%
  mutate(perc_land_use = round(perc_land_use, 2)) %>% 
  group_by(noaa_subba, master_cat) %>%
  mutate(parcel_area = sum(parcel_area)) %>% 
  mutate(perc_land_use = round(sum(perc_land_use), 2)) %>% unique()%>% 
  right_join(subbasins_key)
```

##Land Use Figure
```{r}
#preparing the data for the figure
  #arrange in descending
  #make factors


#figure itself
land_use_barchart<- summary_land_use %>% 
  ggplot(aes(x = noaa_subba, 
             y = parcel_area, fill = master_cat)) +
  theme_minimal() +
  geom_col() +
  labs(fill = "Land Use Category", 
       x = "", 
       y = "Percent of Total Area",
       title = "Primary Land Use Category",
       subtitle = "by subbasin",
       caption = "Data Source: ") +
  coord_flip() +
  # scale_fill_manual(values = custom_colors,
  #                   labels = c("Riparian Planting",
  #                              "Engineered Log Jam",
  #                              "Floodplain"),
  #                   guide = guide_legend(reverse = TRUE)) +
  # geom_text(aes(y = label_y, label = round(n_diff)), color = "white", alpha = 0.5) +
  
  # pushing y axis labels to edhe of data 
  scale_y_continuous(expand = c(0,0)) +
  
  # scale_y_continuous(limits = c(0, 85)) +
  theme(
    #axes
    axis.text = element_text(size = 10),
    axis.title.x = element_text(margin = margin(t = 12, r = 0, b = 30, l = 20), size = 12),
    #legend
    legend.position = c(0.5, -0.175),
    legend.direction = "horizontal",
    legend.title = element_blank(),
    # legend.box.background = element_rect(color = "grey"))
    # titles
    plot.title = element_text(hjust = 0, vjust = 1.8, size = 16),
    plot.subtitle = element_text(hjust = 0, vjust = 2.2),
    plot.caption = element_text(face = "italic", size = 6),
    # gridlines
    panel.grid.major.y  = element_blank(),
    panel.grid.minor.x  = element_blank()
  ) 

land_use_barchart

ggsave("../final_prods/figures/landuse_barchart.jpeg", plot = land_use_barchart, device = "jpeg", width = 12, height = 7, units = "in")

```


## Mapping land use
```{r}
ggplot() + 
  geom_sf(data = land_use_subs, aes(fill = master_cat))
```

#histogram of ???
```{r}
ggplot(flow_stl, aes( x= fpw)) +
  geom_histogram()
```

#################################### Floodplain ###################################
## Creating shapefiles to correspond to the newly created floodplain habitat adjacent to the river

```{r}
#want to create a buffer around flow_stl
bankfull_stl <- st_buffer(flow_stl, dist = flow_stl$bf_width)

#want to create a buffer around the buffer 1/3 * bf_width
floodplain_hab_stl_river<-st_buffer(bankfull_stl, dist = flow_stl$bf_width*(1/3))

st_write(floodplain_hab_stl_river, here("data", "fp_shapefiles", "sidechannel_shapefiles.shp"), append = FALSE)
```

#combine new floodplain shapefile with old floodplain shapefile
```{r}
#changing crs to match floodplain habitat crs
subs_mod <- subs_stl %>% 
  select(noaa_subba) %>% 
  st_transform(26910) 

#filter floodplain to only include historic
hist_flood_stl <- st_join(subs_mod, filter(flood_stl, period %in% "Hist"), join=st_intersects)

#also filter by floodplain habitat with benefit to chinook
flood_ben<-stilly_bens %>% filter(scenario=="Floodplain" &pop == "fall_chinook" & perc_change > 0)  

#filter flood_ben to hist_flood_stl to only get fp with benefits to chinook
hist_flood_stl<-hist_flood_stl %>% filter(noaa_subba %in% unique(flood_ben$subbasin))

# hist_flood_stl<-left_join(flood_ben, hist_flood_stl, by=c("subbasin"="noaa_subba"))
hist_flood_stl<-sf_object <- st_as_sf(hist_flood_stl)

#join marsh/pond floodplain habitat with he habitat adjacent to the river
floodplain_hab_stl <- st_union(hist_flood_stl, floodplain_hab_stl_river)
```

#overlap land use and floodplain shapefile
CHECK THIS ON A MAP
```{r}
#overlap the land use with the floodplain_hab_stl
farm_stl<-land_use_subs %>% filter(master_cat=="Agricultural Area")
floodplain_land_use <- st_join(farm_stl, floodplain_hab_stl, join = st_intersects)
# floodplain_land_use <- st_intersection(farm_stl, floodplain_hab_stl)
```

#plot to see what this looks like (if it is working)
```{r}
ggplot() +
  geom_sf(data = flow_stl, fill = "darkblue") +
  #geom_sf(data = floodplain_land_use, fill = "yellow", color = "yellow") +
  # geom_sf(data = bankfull_stl, fill = "lightblue", color = "black") +
   geom_sf(data = floodplain_land_use, fill = "yellow", color = "yellow") +
  theme_minimal()
```

#filter to just include farmland and multiply by the cost of farmland
```{r}

```

