```{r}
#sourcing data
source("../common.R")

library(sf)
library(rmapshaper)
library(ggplot2)
```


######################################## Land Use #####################################
#Run this only once
```{r}
######################## making shape files to streamline process ##################### 
# 
# # creating a shape file of the unionized stillguamish basin so we don't have to redo this operation every time
# 
# subs_stl_union <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
#   clean_names() %>%
# st_union() %>%
#   st_as_sf() %>%
#   st_write(here("data", "HARP", "Subbasins_STL", "unionized_basin.shp"))
# 
# 
# 
# #creating a shape file of stilly land use so we dont have to read in all the stillaguamish and skagit data and then transform and filter it each time -->
# 
# land_use_stilly <- land_use %>%
# st_intersection(subs_union) %>%
# st_write(here("data", "land_use", "land_use_stilly.shp"))
```

#READING IN THE ABOVE DATA
```{r}
#reading in unionized stilly subasin shape file -->
subs_union <- st_read(here("data", "HARP", "Subbasins_STL", "unionized_basin.shp")) %>% 
  st_transform("EPSG:26910")

#reading in individual stilly subbasins shape file -->
subs_stl <- st_read(here("data", "HARP", "Subbasins_STL", "Subbasins_STL_20220928.shp")) %>%
  clean_names() %>% 
  select(noaa_subba, shape_area, geometry) %>% 
  st_transform("EPSG:26910") 

#land use parcel data for snohomish and skagit county -->
land_use_stilly <- st_read(here("data", "land_use", "land_use_stilly.shp")) %>% 
  clean_names()
```

#NEW PARCELS LAYER BY SUBBASIN 
```{r}
#creating a new df with parcels by subbasin 

#splitting parcels if they intersect w/ multiple subs 
land_use_subs <- st_intersection(land_use_stilly, subs_stl) %>% 
  #joining attributes of noaa subbasin to parcel data 
  st_join(subs_stl, largest = TRUE) %>% 
  select(master_cat, sub_cat, noaa_subba = noaa_subba.x, subba_area = shape_area.x) %>% 
  mutate(parcel_area = st_area(.))

###THIS DOES NOT RUN
# summary <- land_use_subs %>% 
#   group_by(noaa_subba, master_cat) %>% 
#   summarise(perc_land_use = sum(parcel_area)/subba_area*100)
```

## Mapping land use
```{r}
ggplot() + 
  geom_sf(data = land_use_subs, aes(fill = master_cat))
```

#histogram of ???
```{r}
ggplot(flow_stl, aes( x= fpw)) +
  geom_histogram()
```

#################################### Floodplain ###################################
## Creating shapefiles to correspond to the newly created floodplain habitat adjacent to the river

```{r}
#want to create a buffer around flow_stl
bankfull_stl <- st_buffer(flow_stl, dist = flow_stl$bf_width)

#want to create a buffer around the buffer 1/3 * bf_width
floodplain_hab_stl_river<-st_buffer(bankfull_stl, dist = flow_stl$bf_width*(1/3))

```

#combine new floodplain shapefile with old floodplain shapefile
```{r}
#changing crs to match floodplain habitat crs
subs_mod <- subs_stl %>% 
  select(noaa_subba) %>% 
  st_transform(26910) 

#filter floodplain to only include historic
hist_flood_stl <- st_intersection(subs_mod, filter(flood_stl, period %in% "Hist"))

#also filter by floodplain habitat with benefit to chinook
flood_ben<-stilly_bens %>% filter(scenario=="Floodplain" &pop == "fall_chinook" & perc_change > 0)

#left join flood_ben to hist_flood_stl to only get fp with benefits to chinook
hist_flood_stl<-left_join(flood_ben, hist_flood_stl, by=c("subbasin"="noaa_subba"))
hist_flood_stl<-sf_object <- st_as_sf(hist_flood_stl)

#join marsh/pond floodplain habitat with he habitat adjacent to the river
floodplain_hab_stl <- st_union(hist_flood_stl, floodplain_hab_stl_river)
```

#overlap land use and floodplain shapefile
```{r}
#overlap the land use with the floodplain_hab_stl
floodplain_land_use <- st_intersection(land_use_subs, floodplain_hab_stl)
```

#plot to see what this looks like (if it is working)
```{r}
ggplot() +
  geom_sf(data = flow_stl, fill = "darkblue") +
  #geom_sf(data = floodplain_land_use, fill = "yellow", color = "yellow") +
  # geom_sf(data = bankfull_stl, fill = "lightblue", color = "black") +
   geom_sf(data = floodplain_hab_stl, fill = "yellow", color = "yellow") +
  theme_minimal()
```
